### 2018-03-30 kanwal-analysis/BDA-analysis.Rmd
R analysis  -- this helps because we can run the model from R and not worry about cutting off the outputs (as we have running from command line)

---
title: "Kanwal (2017) BDA Notebook"
author: "Ben"
date: "3/30/2018"
output: html_document
---

```{r packages, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(rwebppl)
library(tidyr)
```

```{r echo=FALSE}
# getModelFile()
# ==============
# Return a model string from a .wppl file
#
# Parameters
# ----------
# modelFile: str
#  File path str
# 
# Returns
# -------
# str
#    Model str.
#
getModelFile <- function(modelFile) {
  readChar(modelFile, file.info(modelFile)$size)
}
```

## BDA data
```{r generative-model}
fPath <- "../kanwal-generative-model-20180328.wppl"
m <- getModelFile(fPath)
```

Get posterior predictives...
```{r run-webppl}
dfSynthetic <- rwebppl::webppl(m)
```

## Empirical Kanwal (2017) data
```{r get-empirical-data}
fEmpiricalPath <- "../kanwal-preprocessed.csv" 
dfEmpirical <- read.csv(fEmpiricalPath)
```

```{r prepocessing-synthetic-data-names}
# Minor preprocessing
dfSynthetic <- dfSynthetic %>%
  rename(display_type=trialType) %>%
  mutate(label_type=ifelse(utterance=='zop', 'short', 'long'))
```

## Preprocessing
```{r prepocessing-synthetic-data-counts}
dfSyntheticCnts <-  dfSynthetic %>%
  group_by(IP, condition, display_type, label_type) %>%
  summarise(cnt=n())
```

Fill missing vals synthetic data (note I've alread carried this out in `analysis.ipynb` for empirical data)
```{r prepocessing-synthetic-data-fill-missing-vals}
display_types <- c('frequent', 'infrequent')
label_types <- c('short', 'long')

# New data frame for merging missing counts
dfPlaceholder <- data.frame(expand.grid(IP=unique(dfSynthetic$IP),
                                        condition=unique(dfSynthetic$condition),
                                        display_type=display_types,
                                        label_type=label_types)) %>%
  mutate(
    IP=as.character(IP),
    condition=as.character(condition),
    display_type=as.character(display_type),
    label_type=as.character(label_type),
    cnt=as.integer(0))

# First combine and take counts
dfJoin1 <- full_join(dfSyntheticCnts, dfPlaceholder, by=c('IP', 'condition', 'display_type', 'label_type', 'cnt')) %>%
  arrange(IP, condition, display_type) %>%
  group_by(IP, condition, display_type, label_type) %>%
  summarise(cnt=sum(cnt))

# We want to filter based on whether the group (IP, condition) ash *any* counts, so get these IDs
dfJoin2 <- full_join(dfSyntheticCnts, dfPlaceholder, by=c('IP', 'condition', 'display_type', 'label_type', 'cnt')) %>%
  arrange(IP, condition, display_type) %>%
  group_by(IP, condition, display_type, label_type) %>%
  summarise(cnt=sum(cnt)) %>%
  group_by(IP, condition) %>%
  summarise(groupTotal=sum(cnt))

dfSyntheticFinal <- full_join(dfJoin1, dfJoin2, by=c('IP', 'condition')) %>%
  filter(groupTotal == 32) %>%
  mutate(condition_total=ifelse(display_type=='frequent', 24, 8)) %>%
  group_by(IP, condition, display_type) %>%
  mutate(prop=cnt/condition_total) %>%
  select(IP, condition, display_type, label_type, prop) %>%
  filter(label_type=='short')
  # spread(display_type, prop)
```

Join empirical data and synthetic data
```{r preprocessing-join-empirical-synthetic-data}
dfSyntheticFinal <- dfSyntheticFinal %>%
  select(IP, condition, display_type, prop)

dfEmpiricalProcessed <- dfEmpirical %>%
  mutate(IP=as.character(IP),
         condition=as.character(condition)) %>%
  gather(display_type, prop, c(frequent, infrequent)) %>%
  select(IP, condition, display_type, prop)

dfEmpSynth <- left_join(dfSyntheticFinal, dfEmpiricalProcessed, by=c("IP", "condition", "display_type")) %>%
  gather(data_type, prop, c(prop.y, prop.x)) %>%
  select(IP, condition, display_type, data_type, prop) %>%
  mutate(data_type=ifelse(data_type=='prop.x', 'synthetic', 'empirical'))
```

### Evaluation

How do model predictions look compared to empirical data?
```{r plot-synthetic-vs-empirical}
dfEmpSynth %>%
  spread(display_type, prop) %>%
  ggplot(aes(x=frequent, y=infrequent, col=data_type, group=IP)) +
    geom_point(alpha=0.7, size=2) +
    geom_line(alpha=0.7, col='grey', lty=2) +
    theme_few() +
    facet_grid(~condition)
```

Correlation
```{r correlation-synthetic-vs-empirical}
dfEmpSynthCor <- dfEmpSynth  %>%
  spread(data_type, prop)

paste("R^2:", round(cor(dfEmpSynthCor$empirical, dfEmpSynthCor$synthetic)^2, 3))
```

```{r plot-correlation}
dfEmpSynth %>%
  spread(data_type, prop) %>%
  ggplot(aes(x=synthetic, y=empirical)) +
    geom_jitter(aes(col=condition), alpha=0.5, width=0.01, height=0.01, size=2) +
    geom_smooth(method='lm') +
    ggtitle("Model vs Empirical predictions Kanwal (2017)") +
    theme_few() +
    theme(plot.title = element_text(hjust = 0.5))
```