### 2018-06-17 analysis/data-explore.Rmd
plotting functionality for individual IP behavior -- gotta figure out what the crazy people are doing.

---
title: "data-explore"
author: "Ben"
date: "6/16/2018"
output: html_document
---

```{r libs}
library(ggthemes)
library(tidyverse)
```

```{r options}
options(stringsAsFactors = FALSE)
```

```{r data}
df <- read.csv("../data/df_agg.csv")
```

```{r add-match-label}
df <- df %>%
  mutate(match=label_type_num==display_type_num)
```

```{r get-conditions}
conditions <- unique(df$condition)
```

```{r plotting-helpers}
plot_IP_behavior_match <- function(cond, df, save_file=TRUE) {
  test_img <- df %>%
    filter(condition == cond) %>%
    ggplot(aes(x=as.factor(trial), y=match, col=display_type, group=1)) +
      geom_point(aes(shape=label_type), size=3) +
      geom_line() +
      theme_few() +
      facet_wrap(~IP, ncol=5) +
      ggtitle(paste("Condition:", cond)) +
      theme(text = element_text(size=12),
            axis.text.x = element_text(angle=90, hjust=1, size=6),
            plot.title = element_text(hjust = 0.5))
  if (save_file) ggsave(paste0("img/test-img-", cond, "-20180616.png"), test_img, width=30, height=30)
}

plot_IP_behavior_label_type <- function(cond, df, save_file=TRUE) {
  test_img <- df %>%
    filter(condition == cond) %>%
    ggplot(aes(x=as.factor(trial), y=label_type, col=display_type, group=1)) +
      geom_point(size=3, alpha=0.5) +
      geom_line() +
      theme_few() +
      facet_wrap(~IP+condition, ncol=5) +
      ggtitle(paste("Condition:", cond)) +
      theme(text = element_text(size=24),
            axis.text.x = element_text(angle=90, hjust=1, size=6),
            plot.title = element_text(hjust = 0.5))
  if (save_file) ggsave(paste0("img/label-type-", cond, "-20180616.png"), test_img, width=30, height=30)
}
```

```{r plots}
for (i in 1:length(conditions)) {
  cond <- conditions[i]
  cat(paste("\nCondition:", cond))
  plot_IP_behavior_match(cond, df)
  plot_IP_behavior_label_type(cond, df)
}
```

```{r bad-ips}
plot_IP_behavior_label_type("accuracy", df)
```