### 2019-01-11 bda/analysis/fit_model.Rmd
update human-model comparison to be proportions instead of counts (to mirror Kanwals analysis). think about t statistical test for the correlation coefs?

---
title: "run_model_fitting"
author: "Ben"
date: "11/16/2018"
output: html_document
---

```{r libraries}
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(rwebppl)
library(tidyverse)
```

```{r}
# getModelFile()
# ==============
# Return a model string from a .wppl file
#
# Parameters
# ----------
# modelFile: str
#  File path str
# 
# Returns
# -------
# str
#    Model str.
#
getModelFile <- function(modelFile) {
  readChar(modelFile, file.info(modelFile)$size)
}
```

```{r}
# All data
f_all_data <- "../data/counts_all.csv"
dfAll <- read.csv(f_all_data)
cat("n=", length(dfAll$IP))
```

```{r}
# f_bda_model <- "../models/bda.wppl"
# model_bda <- getModelFile(f_bda_model)
```

```{r}
# problemIPs <- c('142.196.192.44', 
#                 '173.242.228.167',
#                 '104.57.161.213',
#                 '104.11.62.66',
#                 '184.155.134.46',
#                 '67.168.183.190',
#                 '108.210.65.252')
```

```{r}
# pt <- proc.time()
# df_bda <- rwebppl::webppl(model_bda, data=dfAll, data_var='rData', cores=4)
# et <- proc.time() - pt
# cat(paste("Run time: ", et))

df_predictives <- read.csv("../analysis/df_predictives_20181119.csv")
```

```{r merge bda and raw data}
df_predictives <- merge(dfAll, df_predictives)
# write_csv(df_predictives, 'df_predictives_20181119.csv')
```

Check out Model / Human `r^2`
```{r}
df_frequent <- df_predictives %>% 
  select(IP, condition, alpha, depth, inferredCondition, frequent_cnt, posteriorPredictiveFrequent) %>%
  mutate(type='frequent') %>%
  rename(actualCnt=frequent_cnt,
         estCnt=posteriorPredictiveFrequent)

df_infrequent <- df_predictives %>% 
  select(IP, condition, alpha, depth, inferredCondition, infrequent_cnt, posteriorPredictiveInfrequent) %>%
  mutate(type='infrequent') %>%
  rename(actualCnt=infrequent_cnt,
         estCnt=posteriorPredictiveInfrequent)

df_predictives2 <- rbind(df_frequent, df_infrequent)
cor(df_predictives2$actualCnt, df_predictives2$estCnt)^2

```

```{r}
plot_all_post_preds <- df_predictives2 %>%
  # mutate(actualFrequentProps=frequent_cnt / frequent_n,
  #        actualInrequentProps=infrequent_cnt / infrequent_n,
  #        estFrequentProps=posteriorPr5edictiveFrequent / frequent_n,
  #        estInfrequentProps=posteriorPredictiveInfrequent / infrequent_n) %>%
  # select(IP, condition, inferredCondition, actualFrequentProps, estFrequentProps, actualInrequentProps, estInfrequentProps) %>%
  # gather(type, value, c(posteriorPredictiveFrequent, posteriorPredictiveInfrequent))
  ggplot(aes(x=actualCnt, y=estCnt, col=condition), alpha=0.4) +
    geom_smooth(aes(group=0), method='lm') +
    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
    ylab('Posterior predictive \n counts of short form') +
    xlab('Human data \n counts of short form') +
    ggtitle('r^2=0.987') +
    theme_few() +
    theme(plot.title = element_text(hjust = 0.5))

ggsave('../../paper/figs/all-post-predictives-plot.png', plot_all_post_preds, width=6, height=3)
```

## Counts plot 1

```{r}
df_predictives2 %>%
  # mutate(actualFrequentProps=frequent_cnt / frequent_n,
  #        actualInrequentProps=infrequent_cnt / infrequent_n,
  #        estFrequentProps=posteriorPr5edictiveFrequent / frequent_n,
  #        estInfrequentProps=posteriorPredictiveInfrequent / infrequent_n) %>%
  # select(IP, condition, inferredCondition, actualFrequentProps, estFrequentProps, actualInrequentProps, estInfrequentProps) %>%
  # gather(type, value, c(posteriorPredictiveFrequent, posteriorPredictiveInfrequent))
  ggplot(aes(x=actualCnt, y=estCnt, col=condition), alpha=0.4) +
    geom_smooth(method='lm') +
    geom_point(aes(fill=condition), alpha=0.6, size=3, colour="white", pch=21) +
    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
    facet_wrap(~condition, scales="free") +
    ylab('Posterior predictive \n counts of short form') +
    xlab('Human data \n counts of short form') +
    theme_few()

ggsave('../../paper/figs/post-predictives-facet-plot.png', width=6, height=3)
```

## Counts plot 2

Compare to "optimal baseline" optimal in that `time` condition always pick short, `accuracy` always long, etc.
```{r}
bda_vs_baseline_plot <- df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequent', 8,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  gather('model', 'pred', c(baseline, estCnt)) %>%
  mutate(model=ifelse(model=='estCnt', 'BDA', 'Baseline')) %>%
  ggplot(aes(x=actualCnt, y=pred, col=condition), alpha=0.4) +
    geom_point(aes(fill=condition, shape=type), alpha=0.6, size=3) +
    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
    facet_wrap(~model, scales="free") +
    ylab('Baseline model \n counts of short form') +
    xlab('Human data \n counts of short form') +
    theme_few()

ggsave('../../paper/figs/bda-vs-baseline-plot.png', bda_vs_baseline_plot, width=6, height=3)
```

# Proportions Analysis

```{r}
df_cors <- df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequent', 8,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 0,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  gather('model', 'pred', c(baseline, estCnt)) %>%
  group_by(IP, condition, actualCnt, model) %>%
  summarise(cntShortForms=sum(pred)) %>%
  group_by(IP, model, condition) %>%
  summarise(estCnts=sum(cntShortForms),
            actCnts=sum(actualCnt)) %>%
  mutate(condition_n=ifelse(condition=='combined' | condition=='accuracy', 32, 64)) %>%
  mutate(estProp=estCnts/condition_n,
         actProp=actCnts/condition_n) %>%
  ungroup
```

```{r}
df_cors %>%
  group_by(model) %>%
  summarise(n=n(),
            r=cor(estProp, actProp)^2)
```

## Proportions plot1

```{r}
df_cors %>%
  mutate(model=ifelse(model=='baseline', 'Baseline', 'BDA RSA')) %>%
  ggplot(aes(x=actProp, y=estProp, col=condition), alpha=0.4) +
    geom_point(aes(fill=condition), alpha=0.6, size=3) +
    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
    facet_wrap(~model) +
    ylab('Baseline model \n counts of short form') +
    xlab('Human data \n counts of short form') +
    ylim(0, 1) +
    xlim(0, 1) +
    theme_few()

ggsave('../../paper/figs/bda-vs-baseline-plot.png', plot = last_plot(), width=6, height=3)
```

Some kind of significance test?

```{r}
fisher.z <- function(r1,r2,n1,n2) {
  # Todo: review (https://stats.stackexchange.com/questions/64152/are-two-pearson-correlation-coefficients-different)
  ((0.5*log((1+r1)/(1-r1)))-(0.5*log((1+r2)/(1-r2))))/((1/(n1-3))+(1/(n2-3)))^0.5
}

```

```{r}
rs <- df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequent', 8,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  filter(type=='frequent') %>%
  gather(model, value,  c(estCnt, baseline)) %>%
  select(actualCnt, model, value, condition) %>%
  group_by(model) %>%
  summarise(n=n(),
            r=cor(actualCnt, value)^2)
print(rs)
```

## Subject-level inferred params?

```{r}
plot_alpha <- df_predictives %>%
  ggplot(aes(alpha)) +
    geom_density() +
    theme_few()

plot_depth <- df_predictives %>%
  ggplot(aes(depth)) +
    geom_histogram(bins=3) +
    theme_few()

plot_attention <- df_predictives %>%
  ggplot(aes(inferredAttention)) +
    geom_density() +
    theme_few()

plot_condition <- df_predictives %>%
  mutate(inferredCondition=as.character(inferredCondition)) %>%
  mutate(inferredCondition=ifelse(inferredCondition=='conventionalized', 'other', inferredCondition)) %>%
  group_by(inferredCondition) %>%
  summarise(count=n()) %>%
  ggplot(aes(x=inferredCondition, y=count)) +
    geom_bar(stat='identity', position='dodge') +
    theme_few() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_all_params <- grid.arrange(plot_alpha, plot_depth, plot_attention, plot_condition, nrow=2)
ggsave('../../paper/figs/inferred-params-plot.png', plot=plot_all_params, width=6, height=4)
# plot_condition <- df_predictives %>%
#   group_by(inferredCondition) %>%
#   summarise(n=n())
```

How well do we infer the strategies?
```{r}
mean(df_predictives$inferredCondition == df_predictives$actualCondition)
```

We use conventionatlized only for a few. These are very strange in that they *almost* always avoid using the short form for the frequent term, but use it for the non-frequent 
```{r}
df_predictives %>%
  filter(inferredCondition=='conventionalized')
```

Output inferred subject-level params
```{r}
df_predictives %>%
  select(IP, alpha, depth, inferredCondition) %>%
  toJSON() %>%
  write_lines('subjectInferredParams.json')
```

# Try to model time dynamics

```{r}
f_temporal_bda_model <- "../models/bdaFig5.wppl"
model_temporal_bda <- getModelFile(f_temporal_bda_model)
df_trajectories <- rwebppl::webppl(model_temporal_bda)
```

```{r}

df2 <- data.frame()
for (i in seq(1, length(df_trajectories))) {
  df2 <- rbind(df2, df_trajectories[[i]])
}

```
```{r}
numIPs <- length(unique(df2$IP))
df2$trialNum <- rep(seq(1, 31), numIPs)
```

```{r}
df2 %>%
  filter(trialNum==1 | trialNum==31) %>%
  spread(trialNum, data) %>%
  rename('first'=`1`, 'last'=`31`) %>%
  mutate(diff=last-first) %>%
  ggplot(aes(x=IP, y=diff)) +
    geom_bar(stat='identity', position='dodge') +
    theme_few()
```

```{r}
df2 %>%
  group_by(trialNum) %>%
  summarise(n=n(),
            avgProp=mean(data),
            ci_high=avgProp + 1.96 *sqrt(avgProp*(1-avgProp)/n),
            ci_low=avgProp - 1.96 *sqrt(avgProp*(1-avgProp)/n)) %>%
  ggplot(aes(x=trialNum, y=avgProp)) +
    geom_line() +
    geom_errorbar(aes(ymax=ci_high, ymin=ci_low)) +
    ylim(0, 1) +
    theme_few()
```