### 2019-01-23 bda/analysis/fit_model.Rmd
separate model fitting and plotting

---
title: "run_model_fitting"
author: "Ben"
date: "11/16/2018"
output: html_document
---

```{r libraries, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(rwebppl)
library(tidyverse)
```

```{r helpers}
# getModelFile()
# ==============
# Return a model string from a .wppl file
#
# Parameters
# ----------
# modelFile: str
#  File path str
# 
# Returns
# -------
# str
#    Model str.
#
getModelFile <- function(modelFile) {
  readChar(modelFile, file.info(modelFile)$size)
}
```

```{r track_problem_ips}
# Problem IPs saving just to track
# problemIPs <- c('142.196.192.44', 
#                 '173.242.228.167',
#                 '104.57.161.213',
#                 '104.11.62.66',
#                 '184.155.134.46',
#                 '67.168.183.190',
#                 '108.210.65.252')
```

```{r model-run-data-loading}
RUN_MODEL <- FALSE

# User set this
# -------------
MODEL_TYPE <- "NAIVE_BASELINE"
# -------------

# Experimental data we'd like to model
f_kanwal_counts_data <- "../data/counts_all.csv"
# Note this data is restricted to "accuracy", "combined" and "time" sections.
dfAll <- read.csv(f_kanwal_counts_data)
cat("n=", length(dfAll$IP), "\n")

if (RUN_MODEL) {
  pt <- proc.time()
  # Baseline or Full BDA
  model_file <- getModelFileFromType(MODEL_TYPE)
  model_bda <- getModelFile(model_file)
  df_bda <- rwebppl::webppl(model_bda, data=dfAll, data_var='rData', cores=4)
  # Write file
  write.csv(df_bda, file=file.path("../analysis", paste0("df_", MODEL_TYPE, ".csv")))
  et <- proc.time() - pt
  cat(paste("Run time: ", et))
}
```

```{r}
# data2predictions()
# ==================
# Minor pre-processing for correlation analysis.
#
# Parameters
# ----------
# modelFile: str
#  Model type.
# 
# Returns
# -------
# file.path
#    File path to model.
#
data2predictions <- function(modelType) {
  df <- read.csv(getModelFittedModelData(modelType))
  df_frequent <- df %>%
    select(IP, actualCondition, actualFrequent, posteriorPredictiveFrequent, inferredAttention) %>%
    mutate(type='frequent') %>%
    rename(actualCnt=actualFrequent,
           estCnt=posteriorPredictiveFrequent)
  df_infrequent <- df %>%
    select(IP, actualCondition, actualInfrequent, posteriorPredictiveInfrequent, inferredAttention) %>%
    mutate(type='infrequent') %>%
    rename(actualCnt=actualInfrequent,
           estCnt=posteriorPredictiveInfrequent)
  df_combined <- rbind(df_frequent, df_infrequent) %>%
    mutate(modelType=modelType)
  df_combined
}

# deterministicBaselineDf()
# =========================
# Kanwal et al optimal deterministic strategy.
#
# Parameters
# ----------
# df: Data.Frame
#   Data frame of model predictions
# Returns
# -------
# Data.Frame
#    Data frame with deterministic condition-optimal data.
#
deterministicBaselineDf <- function(df) {
  df %>%
    select(IP, actualCondition, actualCnt, type) %>%
    mutate(estCnt=ifelse(actualCondition=='combined' & type=='frequent', 24,
                         # Only use short form for frequent in combined
                         ifelse(actualCondition=='combined' & type=='infrequent', 0,
                                # Never use short form during accuracy
                                ifelse(actualCondition=='accuracy' & type=='frequent', 0,
                                       ifelse(actualCondition=='accuracy' & type=='infrequent', 0,
                                              # Always use short form during time
                                              ifelse(actualCondition=='time' & type=='frequent', 48, 16)))))) %>%
    mutate(modelType='DETERMINISTIC BASELINE',
           inferredAttention=1)
}
```